Accelerating adaptive ultrasound algorithms using graphics processing units \vspace{0.5cm}

### 2014-05-14 ThesisJP/Background/background.tex
Some small final corrections.
Some small corrections.

To "know" the direction of interfering sources the Capon beamformer has to estimate the statistics of the impinging wave field. This is done by estimating a sample covariance matrix, which for an active broadband system can be done in the following way \cite{Synnevag2009}:
\begin{align}
\mat{\breve{R}}[n] &= \frac{1}{N_LN_K}\sum_{n'=n-K}^{n+K} \sum_{l=0}^{N_L-1} \vec{x}_l[n']\vec{x}_l[n']^H\\
\vec{x}_l[n'] &= [x_l[n'], \dotso, x_{l+L}[n']]
\end{align}
where\nomi{$K$}{Temporal averaging parameter (Capon)} $N_K = 2K + 1$ \nomi{$N_K$}{Number of samples used for temporal averaging (Capon)} is the number of samples used for temporal smoothing,\nomi{$L$}{Subarray length (Capon)} $\vec{x}_l$ is a subarray of length $L$, and $N_L = M-L+1$\nomi{$N_L$}{Number of subarrays used for spatial averaging (Capon)} is the number of subarrays used for spatial averaging.  Temporal averaging is used to obtain DAS-like speckle statistics in combination with large subarrays \cite{Synnevag2007a}, and subarray averaging is applied to reduce signal cancellation caused by coherent signals \cite{Reddy1987}. The number of temporal samples is selected relative to the pulse length. To ensure robustness against model errors and numerical stability the matrix is further loaded with a diagonal factor, 
\begin{align}
\epsilon &= d\times\text{trace}\{\mat{\breve{R}}\}/L\\
\mat{\hat{R}} &= \mat{\breve{R}} + \epsilon\mat{I}.
\end{align} 
This corresponds to a white noise component proportional to the in-coherent signal power is added to the estimate \cite{Featherstone1997b}.

@@

The principle behind ray casting is depicted in Fig.\,\ref{fig:vr}. The projection of volumetric data is performed by casting a ray out from every pixel in the output image (Fig.\,\ref{fig:vr_rc}). Along each ray, samples are gathered from the volume and assigned a color and opacity. For ultrasound visualization, the color intensity is dictated by the ultrasound backscatter intensity, and the colorfulness is often encoded with depth information in order to improve depth perception. An example of a volume rendering of cardiac ultrasound data is presented in Fig.\,\ref{fig:dvr}. The opacity, however, is governed by a user-controlled \textit{opacity transfer function} (\nom{OTF}{Opacity transfer function (volume rendering)}) as shown in Fig.\,\ref{fig:vr_otf}. For ultrasound visualization a monotonically increasing function is typically used, which means that below a certain threshold all samples will be rendered transparent (opacity = 0). We will refer to this as the \textit{opacity threshold}, and for the example in Fig.\,\ref{fig:vr_otf} this threshold is 25.

@@

In \textbf{Paper IV} we extend the work by Honigmann \textit{et al.} \cite{Honigmann2003} on a global adaptive OTF by formulating an regional adaptive OTF for visualization of cardiac ultrasound volumes. This regional adaptive OTF is based on estimated statistics of the clutter noise inside the left ventricle (the blood pool). The added complexity of this technique is caused by the statistical estimation, which is basically a second volume rendering (mean and variance projection) that has to be performed before the rendered image is produced.

>>>>

To "know" the direction of interfering sources the Capon beamformer has to estimate the statistics of the impinging wave field. This is done by estimating a sample covariance matrix, which for an active broadband system can be done in the following way \cite{Synnevag2009}:
\begin{align}
\mat{\breve{R}}[n] &= \frac{1}{N_LN_K}\sum_{n'=n-K}^{n+K} \sum_{l=0}^{N_L-1} \vec{x}_l[n']\vec{x}_l[n']^H\\
\vec{x}_l[n'] &= [x_l[n'], \dotso, x_{l+L}[n']]
\end{align}
where\nomi{$K$}{Temporal averaging parameter (Capon)} $N_K = 2K + 1$ \nomi{$N_K$}{Number of samples used for temporal averaging (Capon)} is the number of samples used for temporal smoothing,\nomi{$L$}{Subarray length (Capon)} $\vec{x}_l$ is a subarray of length $L$, and $N_L = M-L+1$\nomi{$N_L$}{Number of subarrays used for spatial averaging (Capon)} is the number of subarrays used for spatial averaging.  Temporal averaging is used to obtain DAS-like speckle statistics in combination with large subarrays \cite{Synnevag2007a}, and subarray averaging is applied to reduce signal cancellation caused by coherent signals \cite{Reddy1987}. The number of temporal samples is selected relative to the pulse length. To ensure robustness against model errors and numerical stability the matrix is further loaded with a diagonal factor, 
\begin{align}
\epsilon &= d\times\text{trace}\{\mat{\breve{R}}\}/L\\
\mat{\hat{R}} &= \mat{\breve{R}} + \epsilon\mat{I}.
\end{align} 
This adds a white noise component proportional to the in-coherent signal power to the estimate \cite{Featherstone1997b}.

@@

The principle behind ray casting is depicted in Fig.\,\ref{fig:vr}. The projection of volumetric data is performed by casting a ray out from every pixel in the output image (Fig.\,\ref{fig:vr_rc}). Along each ray, samples are gathered from the volume and assigned a color and opacity. For ultrasound visualization, the color intensity is dictated by the ultrasound backscatter intensity, and the colorfulness is often encoded with depth information in order to improve depth perception. An example of a volume rendering of cardiac ultrasound data is presented in Fig.\,\ref{fig:dvr}. The opacity, however, is governed by a user-controlled \textit{opacity transfer function} (\nom{OTF}{Opacity transfer function (volume rendering)}) as shown in Fig.\,\ref{fig:vr_otf}. For ultrasound visualization a monotonically increasing function is typically used, which means that below a certain threshold all samples will be rendered transparent (opacity = 0). We will refer to this as the \textit{opacity threshold} (\nom{$\tau$}{Opacity threshold}), and for the example in Fig.\,\ref{fig:vr_otf} this threshold is 25.

@@

In \textbf{Paper IV} we extend the work by Honigmann \textit{et al.} \cite{Honigmann2003} on a global adaptive OTF by formulating a regional adaptive OTF for visualization of cardiac ultrasound volumes. This regional adaptive OTF is based on estimated statistics of the clutter noise inside the left ventricle (the blood pool). The added complexity of this technique is caused by the statistical estimation, which is basically a second volume rendering (mean and variance projection) that has to be performed before the rendered image is produced.
